Designing Qusecs Optimally: Quasi-uniform or Self-similar Layered Materials as Recursively Decomposed Solutions of Geometric Constraint Systems

### 2015-08-23 doc/GMP2015/revision_3/s03_drp.tex
Lots of work on Revision 3, getting close.

@@

\subsection{Canonical DR-Plan}
In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \candrpcomplexity\ algorithm to find the canonical DR-plan of any independent graph. The definition is as follows:

\begin{definition}\label{def:canonical_drplan}
    The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
    (1) it is a DR-plan of $G$;
    (2) children are rigid vertex-maximal proper subgraphs of the parent; and
    (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
\end{definition}

In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight).
% \sidenote{In this section, $G$ is always assumed to be isostatic.}

% This definition gives the canonical DR-plan a surprisingly strong Church-Rosser property.

@@

\begin{theorem}
\label{theorem:canonical_exists_and_is_optimal}
\label{theorem:canonical_is_optimal}
\label{theorem:main}
    A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.
\end{theorem}

@@

Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$, for all nodes $C$ with children $C_1,\ldots,C_N$  retain children nodes according to the following rules:
\begin{enumerate}
   \item If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
   \item If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
\end{enumerate}

This directly satisfies Properties (2) and (3) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrp{G}$ are rigid vertex-maximal proper subgraphs,  which we shorten to {\em clusters}.  To show Property (1) holds (that this constitutes a DR-plan):
for Case 1 above,  since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Case 2 above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.

@@

First, note  that any  DR-plan $R$ without the Property (2) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (2) without  increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$)
is the subgraph of some cluster of $C$.
Thus without loss of generality, we can assume that an optimal DR-plan satisfies Property (2) of a canonical DR-plan.

The proof of optimality of a canonical DR-plan is by induction on its height.  The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges. The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
For the induction step consider a canonical DR-plan $R$ of  height $t+1$ rooted at a node $C$.
Notice that $R$ represents a canonical DR-plan $R(C)$ for the graphs $C$ corresponding to each of its  descendant nodes.
Thus, from the induction hypothesis, we know that the $R(C_i)$ is optimal for $C_i$.

Thus it is sufficient to demonstrate a set of nodes $S$  that must be present in any  DR-plan $R$ for $C$ that satisfies Property (2), including a known optimal one; and furthermore, for any such DR-plan $R$,  either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $R$ has the minimum possible fan-in of 2.

We show the two claims below.
The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (2) must also satisfy Property (3) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (2), including a known optimal one.
The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the {\em same maximal antichain $S$ of descendants $D$ of $C$}. The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment. I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$. Thus any  DR-plan that satisfies Property (2) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$. The two claims complete the proof that every canonical DR-plan is optimal.

@@

% Now we show that this plan is optimal by considering the two cases. Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
 % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs. Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to consider. Either (1) subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.

\medskip\noindent
\vemph{Claim 1:}
% If some pair is intersecting trivially, then in fact all pairs intersect trivially, thus all must be children by defn of canonical drplan
%
A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial, must be children of $C$ in an optimal DR-plan.

@@

% must be optimal for this node.

We prove by contradiction.
Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic. If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$. So, it must be that $U=C$.
\usestwod
However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices. By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible. As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.

@@

\medskip\noindent
\vemph{Claim 2:}
If some pair in the set of child clusters $C_1,\ldots,C_N$ of $C$ has an isostatic (nontrivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of $C$.

To prove Claim 2, notice that if  $C_i \cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i \cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself. Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.
% are potential choices for the optimal DR-plan as they all create equal fan-in (exactly two) at this level.

@@

To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$. Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children. For convenience, we will assume all subgraphs are induced subgraphs of $C$. We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$. The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
% Since
% \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
% the children of $C_i$ will be
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
% and
% \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
% % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
% % the children of this node will be
% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
% % and
% % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
% for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once. At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children. Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for  $N$ levels,  and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
%
% \medskip\noindent
% This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
%
% -- say this at top `proof is by induction on the level of the dr=plan'
\end{proof}

@@

\begin{observation*}\label{lemma:union_intersection}
If $F_i$ and $F_j$ are non-empty isostatic graphs then the following hold: \\
(1) $F_i\cup F_j$ is not trivial;
(2) $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial;
(3) $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic; and
(4) $F_i\cap F_j$ is not underconstrained.
\end{observation*}

@@

\begin{lemma*}\label{lemma:combined_lemma}
Let $C$ be an isostatic node of a canonical DR-plan, with distinct children $C_1,C_2,\ldots, C_m$. Assume $i\ne j$.
Then
\begin{enumerate}
    \item\label{lemma:wc_intersection_is_C}
    $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

@@

\begin{figure*}\centering%
    %
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s}
        \caption{}\label{fig:c2c3ofk33s:a}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.7\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s_candrp}
        \caption{}\label{fig:c2c3ofk33s:b}
    \end{subfigure}%
    %
    \caption{
    (\ref{fig:c2c3ofk33s:a}) A sequence of doublets ($C_2 \times C_3$) intersecting on triangles, where the edges of the triangles are replaced by $K_{3,3}$'s. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows. (\ref{fig:c2c3ofk33s:b}) The canonical DR-plan of $G$, drawn as a DAG. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. Below the third level, the obvious pattern continues until only the individual doublets are present (fourth level) with the ellipses indicating the remaining doublets between those shown. Decomposition of one of these doublets is shown. The dashed lines indicated that this exact decomposition (of the similar nodes on the level) is repeated. Further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges is omitted from the figure.
    }
    \label{fig:c2c3ofk33s}
\end{figure*}%

\begin{example}[DR-plan for self-similar structure]
% \myexample
% \textsl{[DR-plan for self-similar structure]}
This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, the canonical DR-plan of $G$. It begins with the whole (isostatic) graph as the root. The graph $G$ has only two isostatic vertex-maximal subgraphs: $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle (triangle $N$). These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic. As explained in the proof of Theorem \ref{theorem:main}, since there are only 2 possible children, their intersection must be a node 2 levels below the parent. As expected, it is on the third level, as a child of both of $G$'s children.

@@

\subsection{Algorithm}

% \ClearMyMinHeight
% \SetMyMinHeight{.4}{../../img/epsfromtikz/demo_graph}
% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_comdrp}
% \SetMyMinHeight{.3}{../../img/epsfromtikz/demo_graph_candrp}

>>>>

\newcommand{\drpx}[1]{P_{#1}}
\newcommand{\comdrpx}[1]{P_{#1}^{com}}
\newcommand{\candrpx}[1]{P_{#1}^{can}}
\newcommand{\seqdrpx}[1]{P_{#1}^{seq}}

% \newcommand{\componentsx}[1]{CMPNTS(#1)}
% \newcommand{\componentsx}[1]{COMPONENTS(#1)}
% \newcommand{\componentsx}[1]{\text{\emph{COMPONENTS}}(#1)}
\newcommand{\componentsx}[1]{\textsc{Components}(#1)}

\newcommand{\Branch}{Branch}
\newcommand{\branch}{branch}
\newcommand{\branches}{branches}

% \newcommand{\branchx}[1]{\text{\emph{branch}}(#1)}
\newcommand{\branchx}[1]{\textsc{Branch}(#1)}
\newcommand{\branchGePG}{$\branchx{G,e,P_G}$}
\newcommand{\branchLfPL}{$\branchx{L,f,P_L}$}

@@

\subsection{Canonical DR-Plan}
In this section, we define a \dfn{canonical} plan to capture those aspects of an optimal DR-plan that mimic the  uniqueness of a complete DR-plan, and we show that the nonunique aspects do not affect optimality for independent (underconstrained or isostatic) graphs. Furthermore, we give an efficient \ComplexityCanDRP\ algorithm to find an optimal DR-plan of any independent graph.
% The definition is as follows:

In this section and in section~\ref{sec:recomb}, any reference to a graph $G$ without further specification is assumed to be isostatic (i.e.\ well-constrained or $(k,l)$-tight). Furthermore, we only consider unions and intersections of graphs that are induced subgraphs of a single parent graph $G$. In this case unions and intersections are well defined. For example, the union (resp.\ intersection) of $F_1 = (V_1, E_1)$ and $F_2 = (V_2, E_2)$ is the subgraph of $G$ induced by $V_1\cup V_2$ (resp.\ $V_1\cap V_2$).

% When using intersections and unions of graphs, we are always considering subgraphs of a single parent graph. In this case unions and intersections are well defined. For example, the union of $G_1=(V_1, E_1)$ and $G_2=(V_2, E_2)$ is the subgraph $(V_1\cup V_2, E_1\cup E_2)$; similarly, $G_1\cap G_2$ is the subgraph $(V_1\cap V_2, E_1\cap E_2)$.

% \begin{definition}\label{def:canonical_drplan}
%     The \dfn{canonical DR-plan} of a graph $G$ satisfies the following three properties:
%     (1) it is a DR-plan of $G$;
%     (2) children are rigid vertex-maximal proper subgraphs of the parent; and
%     (3) if all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
% \end{definition}

\begin{definition}
[Canonical DR-plan]
\label{def:canonical_drplan}
    A \dfn{canonical DR-plan} is a DR-plan that satisfies the additional two properties:
    \begin{enumerate}
        \item \label{def:canonical_drplan:prop1} Children are rigid vertex-maximal proper subgraphs of the parent.
        \item \label{def:canonical_drplan:prop2} If all pairs of rigid vertex-maximal proper subgraphs intersect trivially then all of them are children, otherwise exactly two that intersect non-trivially are children.
    \end{enumerate}
\end{definition}

@@

\begin{theorem}
[Canonical is optimal]
\label{theorem:canonical_exists_and_is_optimal}
\label{theorem:canonical_is_optimal}
\label{theorem:main}
    A canonical DR-plan exists for a graph $G$ and any canonical DR-plan is optimal if $G$ is independent.
\end{theorem}

@@

    Let $\comdrpx{G}$ be the complete DR-plan of the rigid 2D bar-joint graph $G$.
    % Begin with $\comdrp{G}$ of a rigid 2D bar-joint graph $G$.
    For all nodes $C$ with children $C_1,\ldots,C_N$ retain children nodes according to the following rules:
    \begin{enumerate}[(a)]
       \item \label{thm:main:proof:rule1} If $C_i \cap C_j$ is trivial then retain all $C_1,\ldots,C_N$ as children.
       \item \label{thm:main:proof:rule2} If $C_i \cap C_j$ is rigid then select any two out of $C_1,\ldots,C_N$ as children.
    \end{enumerate}

    This directly satisfies Properties (\ref{def:canonical_drplan:prop1}) and (\ref{def:canonical_drplan:prop2}) of a canonical DR-plan (see Definition~\ref{def:canonical_drplan}), because all the nodes in $\comdrpx{G}$ are rigid vertex-maximal proper subgraphs, which we shorten to \dfn{clusters}. To show that a canonical DR-plan is, in fact, a DR-plan:
    for Rule (\ref{thm:main:proof:rule1}) above, since we start with a complete DR-plan, if we preserve all the children it is still a DR-plan; for Rule (\ref{thm:main:proof:rule2}) above, we know that the union must be rigid as well and it cannot be anything other than $C$, otherwise we would have found a larger rigid proper subgraph of $C$, contradicting vertex-maximality.

@@

    % First, note that any DR-plan $R$ without the Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (\ref{def:canonical_drplan:prop1}) without increasing the max fan-in, since any rigid proper subgraph of a graph $C$ (a child of node $C$ of the DR-plan $R$) is the subgraph of some cluster of $C$.
    First, note that any DR-plan of $G$, $\drpx{G}$, without the Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan can always be modified (by introducing intermediate nodes) to satisfy Property (\ref{def:canonical_drplan:prop1}) without increasing the max fan-in, since any child (a rigid proper subgraph) of node $C$ in $\drpx{G}$  is the subgraph of some cluster of $C$.
    Thus, without loss of generality, we can assume that an optimal DR-plan satisfies Property (\ref{def:canonical_drplan:prop1}) of a canonical DR-plan.

    The proof of optimality of a canonical DR-plan is by induction on its height. The base case trivially holds for canonical DR-plans of height 0, i.e.\ for single edges. The induction hypothesis is that canonical DR-plans of height $t$ are optimal for the root node.
    For the induction step consider a canonical DR-plan $\candrpx{C}$ rooted at node $C$ with height $t+1$.
    Notice that $\candrpx{C}$ contains a canonical DR-plan $\candrpx{C_i}$ for the graphs $C_i$ corresponding to each of $C$'s descendent nodes.
    Thus, from the induction hypothesis, we know that the $\candrpx{C_i}$ is optimal for $C_i$.

    To carry out the induction step, it is sufficient to demonstrate a set of nodes $S$ (of height at most $t$) that must be present in any DR-plan $\drpx{C}$ of graph $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}), including a known optimal one; and furthermore, for any such DR-plan $\drpx{C}$, either (Claim 1) $S$ must be the set of children of $C$; or (Claim 2) for all the ancestors $A$ of $S$, $\drpx{A}$ has the minimum possible fan-in of 2 \todo{Shouldn't it be `node $A$ in $\drpx{C}$ has the minimum...'}.

    We show the two claims below.
    The first claim is that for a node $C$ whose clusters have trivial pairwise intersections, any DR-plan of $C$ that satisfies Property (\ref{def:canonical_drplan:prop1}) must also satisfy Property (\ref{def:canonical_drplan:prop2}) at $C$, i.e.\ the set of children $S$ of $C$ consists of all clusters of $C$.
    Because this is the only choice, it is the minimum fan-in at $C$ for any DR-plan for $C$ with Property (\ref{def:canonical_drplan:prop1}), including a known optimal one.
    The second claim shows that in the case of nodes $C$ whose rigid, vertex-maximal proper subgraphs have  non-trivial pairwise intersections, every canonical DR-plan of $C$ that uses any possible choice of two such subgraphs of $C$ as children results in a minimum possible fan-in of 2 in the ancestor nodes $A$ leading to the \emph{same maximal antichain $S$ of descendants $D$ of $C$}. The antichain is maximal in the partial order of rigid subgraphs of $C$ under containment. I.e.\ $S$ satisfies the property that every proper vertex-maximal rigid subgraph of $C$ is a superset of some $D$ in $S$; this follows from properties of maximal antichains that no element of $S$ is contained in the union of other elements of $S$; and the union of elements of $S$ is $C$. Thus any DR-plan that satisfies Property (\ref{def:canonical_drplan:prop1}) and hence contains two or more of the rigid vertex-maximal proper subgraphs of $C$ as children must also contain every element of $S$. The two claims complete the proof of the induction step and thus the proof that every canonical DR-plan is optimal.

@@

    % Now we show that this plan is optimal by considering the two cases. Observation \ref{lemma:union_intersection} shows that we do not have to consider anything other than the two cases stated in the construction.
     % the intersection of any two isostatic subgraphs can only result in trivial or isostatic subgraphs. Therefore, given $C$ and its isostatic vertex-maximal subgraphs $C_1,\ldots,C_N$, the are only two possibilities to considerthat a canonical DR-plan subgraphs $C_i$ and $C_j$ have a trivial intersection, or (2) they have an isostatic intersection.

    \medskip\noindent
    \vemph{Claim 1:}
    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If the pairwise intersection of clusters is trivial, all of the clusters must be children of $C$ in an optimal DR-plan
    % A set of clusters $C_1,\ldots,C_N$ whose pairwise intersection is trivial must all be children of $C$ in an optimal DR-plan.

@@

    We prove by contradiction.
    % Assume to the contrary that the strict subset $S\subsetneq \{1,\ldots,N\}$ such that $U=\bigcup_{i\in S}{C_i}$ is isostatic. If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the $C_i$. So, it must be that $U=C$.
    % \usestwod
    % However, since $C_i \cap C_j$ is trivial then for $k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more trivial, i.e.\ disconnected vertices. By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible. As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.
    Assume to the contrary that there is a strict subset $S$ of the clusters such that $U$, the union of all elements in $S$, is isostatic. If $U\neq C$, then we found a larger proper subgraph contradicting vertex-maximality of the clusters in set $S$. So, it must be that $U=C$.
    \usestwod
    However, since $C_i \cap C_j$ is trivial then for $C_k\notin S$ we know, by Lemma \ref{lemma:combined_lemma}, Item \ref{lemma:uc_intersection_makes_all_uc}, $U\cap C_k$ must be one or more vertices, i.e.\ disconnected trivial subgraphs. By definition of a DR-plan, $C_k=C\cap C_k$ and we know that $U=C$ so $C_k=U\cap C_k$. Thus, $C_k$ is (i) a collection of disconnected vertices, and (ii) an isostatic subgraph of $C$, which is impossible. As $C$ is isostatic, this means the union of no proper subset of $C_1,\ldots,C_N$ is isostatic, nor is it equal to $C$, proving Claim 1.

@@

    \medskip\noindent
    \vemph{Claim 2:}
    Let the set of clusters of node $C$ be $C_1,\ldots,C_N$. If some pair of clusters
    % If some pair in the set of clusters $C_1,\ldots,C_N$ of $C$
    has an isostatic (non-trivial) intersection, then choosing any two as children (minimum possible fan-in) will result in the same maximal antichain of descendants of node $C$.

    To prove Claim 2, notice that if $C_i\cap C_j$ is isostatic, then, by Observation \ref{lemma:union_intersection}, $C_i\cup C_j$ is also isostatic. This means that, by Lemma \ref{lemma:combined_lemma}, Point \ref{lemma:wc_intersection_makes_all_wc}, the union of any two children of $C$ is $C$ itself. Thus, any two children can be chosen to make a canonical DR-plan and that is the minimum fan-in possible for a node of the DR-plan.

@@

    % To prove this holds, take the set $[N]=\{1,\dots,N\}$, and denote $I=\bigcap_{k\in [N]}{C_k}$ and $R_k=C\setminus C_k$. Suppose  $C_i$ and $C_j$, where $i\neq j$,  are  the children. For convenience, we will assume all subgraphs are induced subgraphs of $C$. We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$. The isostatic vertex-maximal subgraphs of $C_i$ are $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
    Let $I$ denote the intersection of all the clusters; we call this the \dfn{core}. Let $R_i$ be the graph induced by the edge set of $C$ minus the edge set of $C_i$; we call these the \dfn{appendages}.\footnote{\dfn{Core} and \dfn{appendage} are used in Section~\ref{sec:DRP:algo} and are more formally defined in Definition~\ref{def:seqdrp}.} Note that $C$ is the core plus all appendages, and cluster $C_i$ is the core plus all appendages except $R_i$.
    % To prove this holds,
    Suppose $C_i$ and $C_j$, where $i\neq j$, are taken to be the children of node $C$.
    % We know that $C=\induceonc{I\cup\bigcup_{k\in [N]}{R_k}}$ and $C_i=\iunion{i}$.
    The $N-1$ clusters of $C_i$ are the core plus all appendages except $R_i$ and $R_j$, for each $j\neq i$. The pairwise intersection of any of these clusters of $C_i$ will clearly be isostatic, so any two of them are viable children of node $C_i$.
     % $(\iunion{i,1}),\ldots,(\iunion{i,i-1}),(\iunion{i,i+1}),\ldots,(\iunion{i,N})$ all of whose pairwise intersections are isostatic subgraphs. So any two of these are viable children for $C_i$.
    % Since
    % \[C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)\]
    % the children of $C_i$ will be
    % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)\]
    % and
    % \[Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)\]
    % % $C_i=Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i\}}{R_k}\right)$
    % % the children of this node will be
    % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,m\}}{R_k}\right)$
    % % and
    % % $Idc\left(C,I\cup\bigcup_{k\in S_N\setminus\{i,n\}}{R_k}\right)$
    % for arbitrary $m$ and $n$, where $i,j,m,n$ do not equal each other. \todo{Prove these are valid children? Or is this obvious?}
    Beginning with node $C$, this pattern repeats for $N-1$ levels. Every node in this subtree rooted at $C$ has a fan-in of two (the minimum possible) up through this level. At level $N-1$, we have a set of nodes where each node is the core plus some appendage (with every appendage appearing at least once).
    % This continues for $N-1$ levels, always with fan-in of two (the minimum possible), at which point every descendant of $C$ is some $\induceonc{I\cup R_k}$ for $k\in [N]$, with every $k$ appearing at least once.
    % At the last level, there are exactly two rigid proper vertex-maximal subgraphs, and hence a unique choice of pair of children.
    Thus, regardless of the sequence of choices of $C_i$ and $C_j$, and of their descendants at each level, the DR-plan has the optimal fan-in of two for every node for $N-1$ levels, and the collection of last level nodes contain the same maximal antichain of subgraphs (for all choices).
    %
    % \medskip\noindent
    % This proof then applies to itself recursively to show that the fan-in of the children will also be minimum.
    %
    % -- say this at top `proof is by induction on the level of the dr=plan'
\end{proof}

@@

\begin{observation*}
\label{lemma:union_intersection}
    If $F_i$ and $F_j$ are isostatic subgraphs of an independent graph then the following hold:
    % \begin{enumerate}
    %     \item $F_i\cup F_j$ is not trivial.
    %     \item $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial.
    %     \item $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic.
    %     \item $F_i\cap F_j$ is not underconstrained.
    % \end{enumerate}
    (1) $F_i\cup F_j$ is not trivial;
    (2) $F_i\cup F_j$ is underconstrained if and only if $F_i\cap F_j$ is trivial;
    (3) $F_i\cup F_j$ is isostatic if and only if $F_i\cap F_j$ is isostatic; and
    (4) $F_i\cap F_j$ is not underconstrained.
\end{observation*}

@@

\begin{lemma*}
\label{lemma:combined_lemma}
    Let $C$ be an isostatic node of a canonical DR-plan with distinct children $C_1,C_2,\ldots, C_N$. Assume $i\ne j$.
    Then
    \begin{enumerate}
        \item\label{lemma:wc_intersection_is_C}
        $C_i\cup C_j$ is isostatic if and only if $C_i\cup C_j = C$.

@@

\begin{figure*}\centering%
    %
    \begin{subfigure}{0.3\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s}
        \caption{}\label{fig:c2c3ofk33s:a}
    \end{subfigure}%
    %
    \hfill
    \begin{subfigure}{0.7\linewidth}\centering
        \includegraphics[height=\myMinHeight]{../../img/svg/revised_c2c3_of_k33s_candrp}
        \caption{}\label{fig:c2c3ofk33s:b}
    \end{subfigure}%
    %
    \caption{
    (\ref{fig:c2c3ofk33s:a}) A sequence of doublets ($C_2 \times C_3$) intersecting on triangles, where the edges of the triangles are replaced by $K_{3,3}$'s. This pattern continues inwards for a total of $N$ triangles, indicated by the dashed arrows. (\ref{fig:c2c3ofk33s:b}) A canonical DR-plan of $G$, drawn as a DAG. $G\setminus\{a_i,b_i,c_i\}$ is shorthand for $G$ difference those nodes and all of the nodes in the corresponding $K_{3,3}$ subgraphs. Below the third level, the obvious pattern continues until only the individual doublets are present (fourth level) with the ellipses indicating the remaining doublets between those shown. Decomposition of one of these doublets is shown. The dashed lines indicated that this exact decomposition (of the similar nodes on the level) is repeated. Further decomposition of $K_{3,3}$ subgraphs into the separate 9 edges is omitted from the figure.
    }
    \label{fig:c2c3ofk33s}
\end{figure*}%

\begin{example}
[DR-plan for self-similar structure]
    This example details the decomposition of the graph in Figure \ref{fig:c2c3ofk33s}, a canonical DR-plan of $G$. It begins with the whole (isostatic) graph as the root. The graph $G$ has only two isostatic vertex-maximal subgraphs: $G$ without the outermost triangle composed of $K_{3,3}$ graphs (triangle $1$) and $G$ without the inner triangle (triangle $N$). These intersect on $G$ without triangle $1$ and $N$ which is clearly isostatic. As explained in the proof of Theorem \ref{theorem:main}, since there are only 2 possible children, their intersection must be a node 2 levels below the parent. As expected, it is on the third level, as a child of both of $G$'s children.

@@

\subsection{Algorithm}
\label{sec:DRP:algo}

The algorithm for finding an optimal DR-plan relies on key structural properties of canonical DR-plans that are revealed by the proof of Theorem~\ref{theorem:canonical_exists_and_is_optimal}. We list these properties before describing the algorithm. Furthermore, in order to deal with the apparent non-uniqueness of a canonical DR-plan (in the case when the children of a node have non-trivial intersections), and to make the canonical DR-plan more malleable for algorithmic purposes, we define a key notion of a \dfn{sequential DR-plan} that is obtained from a canonical DR-plan, and also turns out to be an optimal DR-plan. A sequential DR-plan further clarifies the essential uniqueness of canonical DR-plans. First, we present a recursive definition of the canonical DR-plan, which is more useful in this section, and make an important observation.

\begin{definition}
[Canonical DR-plan]
\label{def:canonical_drp_rec}
    A \dfn{canonical DR-plan} of $G$ is recursively defined as follows:
    \begin{enumerate}
        \item Base case: When $G$ is a single edge, the canonical DR-plan for $G$ is $G$ itself.
        \item In case the pairwise intersections of the proper vertex-maximal rigid subgraphs $C_i$ of $G$  are all trivial, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$.
        \item In case there are two proper vertex-maximal rigid subgraphs $C_i$ and $C_j$ of $G$ with non-trivial intersection, take the children of $G$ to be the roots of the canonical DR-plans for $C_i$ and $C_j$.
    \end{enumerate}
    % When $G$ is a single edge (base case), the canonical DR-plan for $G$ is $G$ itself.
\end{definition}
\todo{What if input is not isostatic? What roots?}

\begin{observation}
[Canonical DR-plan size]
    While all canonical DR-plans of independent graphs are optimal, based on the choice of children in Case (3) above they can have varying \dfn{size} (i.e.\ the number of unique nodes). If we want the canonical DR-plan to have small size (linear \todo{or just polynomial?} in the number of vertices) we can choose the children in the following fashion: In Case (3) above, a canonical DR-plan for $C_i$ (resp.\ $C_j$) is further recursively defined by taking the children of $C_i$ (resp.\ $C_j$) to be the roots of a canonical DR-plan of $C_i\cap C_j$ (which is rigid, by Lemma~\ref{lemma:combined_lemma}), and a canonical DR-plan of $C_i\setminus C_j$ (resp. $C_j\setminus C_i$) (which are not rigid, by Observation~\ref{lemma:union_intersection} and Lemma~\ref{lemma:combined_lemma}).

    Note that a canonical DR-plan can have more than one node representing the same rigid subgraph if and only if Case (3) occurs.
\end{observation}
